### 2015-10-23 coverletter/cover.tex
final fixes

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plain Cover Letter
% LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%       PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{letter} % Default font size of the document, change to 10pt to fit more text
\usepackage{graphicx}
%\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
%\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font

% Margins
\usepackage[left=1.25in,right=1.25in,top=1.5in,bottom=1.25in]{geometry}
%\let\raggedleft\raggedright % Pushes the date (at the top) to the left, comment this line to have the date on the right

\usepackage{eso-pic,graphicx}
 \begin{document}

%----------------------------------------------------------------------------------------
%       ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{Professor A.C. Klein\\
Editor, Nuclear Technology}
%----------------------------------------------------------------------------------------
%       YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------

\address{Robert W. Carlsen\\
229 N Midvale Blvd Apt 1\\
Madison, WI 53705}

%----------------------------------------------------------------------------------------
%       LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------

\opening{Professor Klein,}

Please find enclosed a manuscript entitled: ``Challenging Fuel Cycle Modeling
Assumptions: Facility and Time Step Discretization Effects'' which I am
submitting for exclusive consideration of publication as a research article in
Nuclear Technology.  This manuscript describes work investigating impacts of
common fuel cycle simulator modeling choices in order to enable more
deliberate simulation design and help improve fuel cycle analysis integrity
generally.

Thank you for your consideration of this work. Please address all
correspondence concerning this manuscript to me and feel free to correspond
with me by e-mail.

\closing{
Happy Halloween,

Robert Carlsen
}

%----------------------------------------------------------------------------------------

\end{letter}

\end{document}